Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-17 paper/paper.tex
thorough proofread 1 (sections 1-3) * renamed section dropout -> Variational Dropout * fixed real-ard kl-div eq (8), cleaned out dubious and imprecise statements (em-algo) * removed excessive citations and repeated rewordings in intro and VB * improved language and coherence in Bayes-related narrative (VI, VB, estimator etc.), reworded and rearranged paragraphs for clearer exposition

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and arithmetic (computational)
complexity restrictions \citep{he_amc:_2018}. Therefore, compression methods for achieving
high model sparsity and numerical efficiency without losing much in performance are
especially relevant.
% Solving the storage and artihmetic constraints encourages development of model compression
% and sparsification methods, that focus on favourable trade-off between performance and size.

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind the adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that the key merit of $\cplx$-valued neural
networks is simultaneous phase rotation and amplitude adjustment, intrinsic to multiplication
in $\cplx$ field. Despite inherently higher arithmetic complexity, which stems from up to four
floating point operations per single $\cplx$ multiplication \citep{monning_evaluation_2018}
$\cplx$-arithmetic reduces ineffective degrees of freedom in comparison with an analogous
double-dimensional $\real$-valued network. \citet{hirose_complex-valued_2009} lists numerous
applications related to acoustic and radio signal processing and demonstrates superiority of
$\cplx$-networks in the task of shallow landmine detection using ground penetrating radar
imaging. More recently they have been applied to music transcription and speech recognition
\citep{trabelsi_deep_2017}. \citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs
to deal with exploding~/~vanishing gradients and apply the recurrent models to long-term
dependence prediction tasks. \citet{wisdom_full-capacity_2016} expand on their work by optimizing
recurrent weights on the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018}
investigate different $\cplx$-valued gating mechanisms for unitary-RNNs and conduct experiments
on human motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification.

@@

% on dropout
Dropout \citep{hinton_improving_2012} prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
have shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated multitask learning, then by sampling uncorrelated weight masks dropout acts
as a superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Bayesian inference is a very rich field with a large body of research and techniques that
enable practical quantification of predictive, parameter, or model uncertainty and its 
propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike Bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Bayesian methods with suitable modelling and prior assumptions, such
as Variational Dropout, can be used towards model sparsification
\citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$ \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although variational dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that, despite dropout showing more variable
performance than magnitude pruning, it induces non-uniform sparsity throughout the model,
which has been shown by \citet{he_amc:_2018} to be essential for superior compression.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational and stochastic sparsification techniques we extend Bayesian dropout
for $\cplx$-valued networks. We assess the performance-compression trade-off of the proposed
Bayesian technique by conducting a large-scale numerical study of $\cplx$-valued networks
on two tasks: image recognition on MNIST-like and CIFAR10 datasets and music transcription
on MusicNet \citep{thickstun_learning_2017}. We reproduce the state-of-the-art result by
\citet{trabelsi_deep_2017} for $\cplx$-valued networks on the latter with $\approx 5\%$
of remaining non-zero parameters.

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate, compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes.

@@

The posterior distribution itself is intractable, save for the relatively simple cases, and
therefore exact inference is traded for tractability and scalability offered by Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the Bayes rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and have either tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$ \citep{williams_simple_1992} or can be represented as differentiable transformations
of non-parametric noise \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
%
The natural choice for $\rho$ is Kullback-Leibler divergence of density $q$ from $p$,
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
the main reason being that it satisfies the following identity for well-behaved densities:
% $(q, p, \pi)$
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$ \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citet{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which
makes it possible to employ stochastic gradient-based optimization methods to solve
\eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $\varepsilon_i \sim p_\varepsilon$.
%
% (DSVI) titsias-lazaro.tex \todo{DSVI} \citep{titsias_doubly_2014}

\subsection{Dropout} % (fold)
\label{sub:dropout}

With a special family of posterior approximation, variational inference can be used as
a regularization method and as a model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect \citep{wan_regularization_2013},
and Gaussian dropout \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and propose \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. The divergence term in \eqref{eq:elbo} unravels into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression over $\alpha \in (0, 1)$. \citet{molchanov_variational_2017} refine the
approximation by using a nonlinear regression based on sigmoid and soft-plus.
% see p.6 sec 3.4 of kingma_variational_2015
We independently verify their approximation against Monte-Carlo estimates for $\alpha$
varying over a fine logarithmic grid and the exact expression for the derivative of
\eqref{eq:improper-kl-div-real}, which we derive in appendix~\ref{sub:real-chisq-grad}.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates
%
% (in) it is in poor taste to brag about this begin the first time anyone has derived a
% derivative of an intractable function.

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus, \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another Bayesian dropout
method, called \textit{Automatic Relevance Determination}, has been developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
ARD Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes},
fits a prior distribution while performing variational inference. Maximizing \eqref{eq:elbo}
over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

>>>>

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and arithmetic complexity
restrictions \citep{he_amc:_2018}. Therefore, compression methods for achieving high
model sparsity and numerical efficiency without losing much in performance are especially
relevant.
% Solving the storage and artihmetic constraints encourages development of model compression
% and sparsification methods, that focus on favourable trade-off between performance and size.

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind the adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that the key merit of $\cplx$-valued neural
networks is simultaneous phase rotation and amplitude adjustment, intrinsic to multiplication
in $\cplx$ field. Despite inherently higher arithmetic complexity, which stems from up to four
floating point operations per single $\cplx$ multiplication
$\cplx$-arithmetic reduces ineffective degrees of freedom in comparison with an analogous
double-dimensional $\real$-valued network. \citet{hirose_complex-valued_2009} lists numerous
applications related to acoustic and radio signal processing and demonstrates superiority of
$\cplx$-networks in the task of shallow landmine detection using ground penetrating radar
imaging. More recently they have been applied to music transcription and speech recognition
\citep{trabelsi_deep_2017}. \citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs
to deal with exploding~/~vanishing gradients and apply the recurrent models to long-term
dependence prediction tasks. \citet{wisdom_full-capacity_2016} expand on their work by optimizing
recurrent weights on the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018}
investigate different $\cplx$-valued gating mechanisms for unitary-RNNs and conduct experiments
on human motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification.

@@

% segue into bayesian inference
Bayesian Inference is a principled framework of reasoning about uncertainty, which updates
prior beliefs in accordance with the likelihood of empirical observations into posterior
belief. Under suitable modelling and distributional assumptions, Bayesian methods can be
used towards model sparsification, e.g. Variational Dropout \citep{kingma_variational_2015,molchanov_variational_2017}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare Variational Dropout, magnitude pruning, and $\ell_0$
regularization \citep{louizos_learning_2017} on large-scale models. The latter is a
probabilistic regularization method, which penalizes the expected number of non-zero
entries in $[0, 1]$-valued stochastic parameter masks with learnable hard-concrete
distribution with atoms at $0$ and $1$ \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that Variational Dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ in deep architectures, although pruning is prefered for
simplicity, stability and speed. They also observe that dropout induces non-uniform sparsity
throughout the model, which \citet{he_amc:_2018} has shown to be essential for superior
compression.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational and stochastic sparsification techniques we extend Bayesian dropout
to $\cplx$-valued networks. We assess the performance-compression trade-off of the proposed
Bayesian technique by conducting a large-scale numerical study of $\cplx$-valued networks
on two tasks: image recognition on MNIST-like and CIFAR10 datasets and music transcription
on MusicNet \citep{thickstun_learning_2017}.

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive the penalty terms in the evidence lower bound. In sec.~\ref{sec:experiments}
we evaluate the sparsification rate, compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes.

@@

Save for the relatively simple cases, the posterior distribution is intractable, and
therefore exactness Bayesian Inference is traded for tractability and scalability of
Variational Bayesian Inference. Instead of deriving the posterior $p(\omega \mid D)$,
this approach postulates an optimization problem for finding an approximation $q$,
which is close to $p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The variational approximation class
$\mathcal{Q}$ is chosen so that its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and have either tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$ or can be represented as differentiable transformations of non-parametric noise
\citep{williams_simple_1992,kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradient) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
%
The most common choice for $\rho$ is Kullback-Leibler divergence of density $q$ from $p$,
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
the main reason being that it satisfies the following identity for well-behaved densities:
% $(q, p, \pi)$
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This leads to an equivalent problem for \eqref{eq:variational-progam} of maximizing
the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where $\lambda$, $\theta$ and $\phi$ are optimization parameters of the prior $\pi$,
approximation $q$ and the likelihood $p(D\mid \omega)$, respectively. Other objectives
are possible, besides \eqref{eq:elbo_general}, provided $p(\omega \mid D)$ is evaluated
only through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citet{kingma_auto-encoding_2014},
replaces the Variational objective \eqref{eq:elbo_general} by its unbiased differentiable
Monte-Carlo estimator, which makes it possible to employ stochastic gradient optimization
methods to solve \eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some differentiable $
  g(\varepsilon; \theta)
$ and non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$, then the SGVB objective is
\begin{equation}  \label{eq:sgvb_objective}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$, the KL-divergence term replaced by unbiased
differentiable finite sample estimator if it is intractable.
%
For the purposes of this study, we assume that the likelihood satisfies $
  \log p_\phi(D\mid \omega)
    = \sum_i \log p_\phi(z_i \mid \omega)
$, i.e. the observed data $
  D = (z_i)_{i=1}^N
$ is conditionally independent, and consider \eqref{eq:sgvb_objective} with $L=1$
and random mini-batches from $D$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{k=1}^M
        \log p_{\phi}(z_{i_k} \mid g(\varepsilon_k; \theta))
    \,,
\end{equation}
for a random subsample $(z_{i_k})_{k=1}^M$ and iid $\varepsilon_k \sim p_\varepsilon$.
%
% (DSVI) titsias-lazaro.tex \todo{DSVI} \citep{titsias_doubly_2014}

\subsection{Variational Dropout} % (fold)
\label{sub:variational_dropout}

With a special family of posterior approximation, Variational Inference can be used as
a regularization method and as a model sparsification technique.

% on dropout
Dropout \citep{hinton_improving_2012} prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In a recent study concerning multitask learning
\citet{cheung_superposition_2019} have shown the possibility of storing task-specific
parameters in non-destructive superposition within a single network. In particular, their
argument implies that if the single task setting is viewed as multitask learning with
replicated task, then by sampling uncorrelated binary masks Dropout acts as a superposition
method, utilizing the learning capacity of the network better.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

Dropout can be viewed as a Bayesian model averaging method, where models share parameters
and are weighted equally. Thus, \citet{kingma_variational_2015} consider Dropout, DropConnect
\citep{wan_regularization_2013}, and Gaussian dropout \citep{wang_fast_2013} through the
lens of Bayesian Inference methods and propose \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.

% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. The divergence term in \eqref{eq:elbo} unravels into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression over $\alpha \in (0, 1)$. \citet{molchanov_variational_2017} refine the
approximation by using a nonlinear regression based on sigmoid and soft-plus.
% see p.6 sec 3.4 of kingma_variational_2015
In appendix~\ref{sub:real-chisq-grad} we verify the value and the derivative of their
approximation against the Monte-Carlo estimate of \eqref{eq:improper-kl-div-real} for $\alpha$
varying over a fine logarithmic grid and the gradient of \eqref{eq:improper-kl-div-real},
for which we obtain an exact expression.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates
%
% (in) it is in poor taste to brag about this begin the first time anyone has derived a
% derivative of an intractable function.
The key role in the factorization is played by $\alpha_j$, which reflects the relevance
of the parameter $\omega_j$ it is associated to, by being the ratio of its squared mean
to its variance. Thus, \citet{molchanov_variational_2017} focus on the model sparsification
capabilities of Variational Dropout, optimizing $\alpha$ for each individual parameter.
\citet{kharitonov_variational_2018} address theoretical issues with improper priors and
propose an alternative Bayesian Dropout method based on \textit{Automatic Relevance
Determination} procedure. In essence, the log-uniform factor $\pi(\omega_{ij})$ is replaced by
a proper ARD Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes},
fits a prior distribution while performing Variational Inference. Maximizing \eqref{eq:elbo}
over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \tfrac12 \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}
Flexibility in specification of $q_\theta$ enables developing structured dropout with
shared $\alpha$ across groups of parameters $\omega$.
